### 2022-08-18 doc/src/AdvPLIC.tex
Remove term 'master', replace with appropriate term
I chose 'main' and 'controller' as replacements for the term 'master' as
the situation warranted.

Signed-off-by: Stephano Cetola <scetola@linuxfoundation.org>

\textbf{%
Warning!
This draft specification is likely to change before being accepted as
standard by the {\RISCV} International Association.%
}
\bigskip

In a {\RISCV} system, a Platform-Level Interrupt Controller (PLIC)
handles external interrupts that are signaled through wires rather than
by MSIs.
When the {\RISCV} harts in a system do not have IMSICs, the harts
themselves do not support MSIs, and all external interrupts to such
harts must pass through a PLIC.
But even in machines where harts have IMSICs and most interrupts are
communicated via MSIs, it is not unusual for some device interrupts
still to be signaled by dedicated wires.
In particular, for devices (or device controllers) that do not
otherwise need to be bus masters in the system, the cost of supporting
MSIs is especially high, so wired interrupts are a frugal alternative.
Wired interrupts also continue to be universally supported by all
current computer platforms, unlike MSIs, making another reason for many
commodity devices or controllers to choose wired interrupts over MSIs,
when not implementing a standard like PCI Express that dictates MSIs.

@@

\begin{figure}[th]
\centerline{\includegraphics[scale=0.55]{AdvPLIC-ex-3Domains.png}}
\caption{%
A {\RISCV} system that extends the example of
Figure~\ref{fig:AdvPLIC-ex-2Domains} with a fifth \mbox{M-mode}-only
``master'' hart, with a separate machine-level interrupt domain above
the other domains.%
}
\label{fig:AdvPLIC-ex-3Domains}
\end{figure}

>>>>

\textbf{%
This chapter is considered stable but could still change before being
ratified as standard by the {\RISCV} International Association.%
}
\bigskip

In a {\RISCV} system, a Platform-Level Interrupt Controller (PLIC)
handles external interrupts that are signaled through wires rather than
by MSIs.
When the {\RISCV} harts in a system do not have IMSICs, the harts
themselves do not support MSIs, and all external interrupts to such
harts must pass through a PLIC.
But even in machines where harts have IMSICs and most interrupts are
communicated via MSIs, it is not unusual for some device interrupts
still to be signaled by dedicated wires.
In particular, for devices (or device controllers) that do not
otherwise need to be bus controllers in the system, the cost of supporting
MSIs is especially high, so wired interrupts are a frugal alternative.
Wired interrupts also continue to be universally supported by all
current computer platforms, unlike MSIs, making another reason for many
commodity devices or controllers to choose wired interrupts over MSIs,
when not implementing a standard like PCI Express that dictates MSIs.

@@

\begin{figure}[th]
\centerline{\includegraphics[scale=0.55]{AdvPLIC-ex-3Domains.png}}
\caption{%
A {\RISCV} system that extends the example of
Figure~\ref{fig:AdvPLIC-ex-2Domains} with a fifth \mbox{M-mode}-only
``main'' hart, with a separate machine-level interrupt domain above
the other domains.%
}
\label{fig:AdvPLIC-ex-3Domains}
\end{figure}

### 2022-08-18 doc/src/IOMMU.tex
Add preface and update status of each chapter
Remove term 'master', replace with appropriate term
I chose 'main' and 'controller' as replacements for the term 'master' as
the situation warranted.

Signed-off-by: Stephano Cetola <scetola@linuxfoundation.org>

\textbf{%
Warning!
This draft specification is likely to change before being accepted as
standard by the {\RISCV} International Association.%
}
\bigskip

@@

If a single physical I/O device can be subdivided for multiple masters,
each sub-device is referred to here as one device.

@@

\item
For each I/O device connected to the system through an \mbox{IOMMU},
software can configure at the \mbox{IOMMU} a \emph{device context}, which
associates with the device a specific virtual address space and any
other per-device parameters the \mbox{IOMMU} may support.
By giving devices each their own separate device context at an \mbox{IOMMU},
each device can be individually configured for a different software
master, usually a guest OS or the main (host) OS.
On every memory access made from a device, hardware indicates to
the \mbox{IOMMU} the originating device by some form of unique device
identifier, which the \mbox{IOMMU} uses to locate the appropriate device
context within data structures supplied by software.
For PCI, for example, the originating device may be identified by the
unique triple of PCI bus number, device number, and function number.

>>>>

\textbf{%
Warning!
This chapter is only a draft, and might change significantly before
being ratified as standard by the {\RISCV} International Association.%
}
\bigskip

@@

If a single physical I/O device can be subdivided for multiple controllers,
each sub-device is referred to here as one device.

@@

\item
For each I/O device connected to the system through an \mbox{IOMMU},
software can configure at the \mbox{IOMMU} a \emph{device context}, which
associates with the device a specific virtual address space and any
other per-device parameters the \mbox{IOMMU} may support.
By giving devices each their own separate device context at an \mbox{IOMMU},
each device can be individually configured for a different software
controller, usually a guest OS or the main (host) OS.
On every memory access made from a device, hardware indicates to
the \mbox{IOMMU} the originating device by some form of unique device
identifier, which the \mbox{IOMMU} uses to locate the appropriate device
context within data structures supplied by software.
For PCI, for example, the originating device may be identified by the
unique triple of PCI bus number, device number, and function number.
